|

### 2024-10-15 posts/BarthPapageorgeThom/BPT.qmd
Completed BPT.html for RG2024/10/15

* Gene-wealth gradient exists: EA score predicts wealth at retirement
* Gene-wealth gradient&lArr; education, income, bequests, parents’ education, but does not explain fully
* Gene-wealth gradient&rarr;0 when stock ownership is added
   * EA score $\propto$ risk torelance, longevity, stock ownership, precision in macroeconomic risk perception
* HHs without DB pension: gene-wealth gradient&uarr;

@@

![Chromosone...DNA](DNAMacrostructure.jpg)

@@

::::{.fragment}
![DNA structure](DNAStructure.jpg)

@@

* PGS is considered to be a predictor of individual fixed effects
   * HRS sample: covariance with EA score / total schooling years variance = 10.6%
   * Edu attainment SNPs $\propto$ biological process of brain development, cognition [@Okbay2016; @Lee2018]
   * Edu attainment SNPs $\propto$ cognition SNPs [@Okbay2016]

@@

1. Into genotyping: Older, more educated, more females, wealthier [(Table S1)](BarthPapgergeThome_SNPWealth_JPE2020Appendix.pdf#page=5) 
   * Education $\propto$ genotyping &rArr; individuals with low EA scores have higher than average education  &rArr; attenuates EA score $\propto$ education gradient 
1. Of using retired households: More precise wealth but smaller sample size/more selection [(Table 3)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=17) 
   * Selection results in more males, older in high EA scores &lArr; (Male) higher EA scorers live longer, but modest in size (Panel A)
   * Age band of retired HHs &uarr; than main sample &rArr; q4-q1 differences (more selection) &uarr; &rArr; EA score $\propto$ education gradient &uarr; (Panel A vs. B)
   * $+$non-retired HHs &rArr; q4-q1 differences &darr; (less selection) &rArr; EA score $\propto$ education gradient &darr; (Panel A vs. C)
   * Smaller degree of selection leads to smaller q4-q1 differences
   * But larger measurement errors in wealth ($\because$ pensions, etc. are not measured)

@@

::: {.column width="5%"}
:::{style="font-size: 45%;line-height: 1.1;"}
Fig 2A
:::
:::
::: {.column width="25%"}
![](Fig2A.jpg){width="70%"}
:::
::: {.column width="70%"}
EA score $\propto$ wealth

@@

:::{.fragment}
Table 2 also shows EA score $\propto$ wealth ($475K, q4-q1), lifetime labor income ($380K, q4-q1)  
:::
:::

@@

::: {.column width="70%"}
:::{.fragment}
EA score $\propto$ wealth similar between high schoolers vs college grads, up to q3
:::
:::

@@

::: {.column width="70%"}
:::{.fragment}
EA score $\propto$ wealth gradient: .246 (raw)&rarr;.070 (+edu)&rarr;.047 (+labor income)
:::

@@

::: {.column width="60%"}
* Stock ownership&uarr; but not of other assets 
* EA score $\propto$ wealth gradient &rarr;0 if we add stock ownership  
:::

@@

::: {.column width="60%"}
* EA score $\propto$ wealth, conditional on DB pension participation
* EA score $\propto$ wealth&uarr; if not participating in DB pension
:::

@@

* Previous studies: EA scores $\propto$ education
* This study: EA scores $\propto$ wealth
* EA scores $\propto$ probabilistic thinking &rArr; human capital accumulation heterogeneity  
   * Teaching probability may make human capital more homogenous?
   * Not necessarily, gene-wealth gradient is a biased estimate &larr; environment
   * But it hints something can be done there

>>>>

* Gene-wealth gradient exists: EA score predicts wealth at retirement
* Gene-wealth gradient&lArr; education, income, bequests, parents’ education, but does not explain fully
* Gene-wealth gradient&rarr;0 when stock ownership is added
   * EA score $\SIM$ risk torelance, longevity, stock ownership, precision in macroeconomic risk perception
* HHs without DB pension: gene-wealth gradient&uarr;

@@

![](DNAMacrostructure.jpg){height="100%"}

@@

::::{.fragment}
![](DNAStructure.jpg){height="80%"}

@@

* PGS is considered to be a predictor of individual fixed effects
   * HRS sample: covariance with EA score / total schooling years variance = 10.6%
   * Edu attainment SNPs $\SIM$ biological process of brain development, cognition [@Okbay2016; @Lee2018]
   * Edu attainment SNPs $\SIM$ cognition SNPs [@Okbay2016]

@@

1. Into genotyping: Older, more educated, more females, wealthier [(Table S1)](BarthPapgergeThome_SNPWealth_JPE2020Appendix.pdf#page=5) 
   * Education $\SIM$ genotyping &rArr; individuals with low EA scores have higher than average education  &rArr; attenuates EA score $\SIM$ education gradient

2. Of using retired households: More precise wealth but smaller sample size/more selection [(Table 3)](BarthPapgergeThome_SNPWealth_JPE2020.pdf#page=17) 
   * Selection results in more males, older in high EA scores &lArr; (Male) higher EA scorers live longer, but modest in size (Panel A)
   * Age band of retired HHs &uarr; than main sample &rArr; q4-q1 differences (more selection) &uarr; &rArr; EA score $\SIM$ education gradient &uarr; (Panel A vs. B)
   * $+$non-retired HHs &rArr; q4-q1 differences &darr; (less selection) &rArr; EA score $\SIM$ education gradient &darr; (Panel A vs. C)
   * Smaller degree of selection leads to smaller q4-q1 differences
   * But larger measurement errors in wealth ($\because$ pensions, etc. are not measured)

@@

::: {.column width="5%"}
:::{style="font-size: 45%;line-height: 1.1;"}
Fig 2A
:::
:::
::: {.column width="25%"}
![](Fig2A.jpg){width="70%"}
:::
::: {.column width="70%"}
EA score $\SIM$ wealth

@@

:::{.fragment}
Table 2 also shows EA score $\SIM$ wealth ($475K, q4-q1), lifetime labor income ($380K, q4-q1)  
:::
:::

@@

::: {.column width="70%"}
:::{.fragment}
EA score $\SIM$ wealth similar between high schoolers vs college grads, up to q3
:::
:::

@@

::: {.column width="70%"}
:::{.fragment}
EA score $\SIM$ wealth gradient: .246 (raw)&rarr;.070 (+edu)&rarr;.047 (+labor income)
:::

@@

::: {.column width="60%"}
* Stock ownership&uarr; but not of other assets 
* EA score $\SIM$ wealth gradient &rarr;0 if we add stock ownership  
:::

@@

::: {.column width="60%"}
* EA score $\SIM$ wealth, conditional on DB pension participation
* EA score $\SIM$ wealth&uarr; if not participating in DB pension
:::

@@

* Previous studies: EA scores $\SIM$ education
* This study: EA scores $\SIM$ wealth
* EA scores $\SIM$ probabilistic thinking &rArr; human capital accumulation heterogeneity  
   * Teaching probability may make human capital more homogenous?
   * Not necessarily, gene-wealth gradient is a biased estimate &larr; environment
   * But it hints something can be done there
